---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(lubridate)
library(gganimate)
theme_set(theme_minimal())
```


```{r}
bird_counts <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-06-18/bird_counts.csv",
                               col_types = cols(
  year = col_integer(),
  species = col_character(),
  species_latin = col_character(),
  how_many_counted = col_double(),
  total_hours = col_double(),
  how_many_counted_by_hour = col_double()
)
                               )

bird_counts %>% 
  head(100)
```


```{r}
bird_counts %>% 
  skimr::skim()
```

So there is some missing data.

```{r}
bird_counts %>% 
  group_by(species) %>% 
  summarise_at(vars(how_many_counted, total_hours), ~sum(.x, na.rm = T))
  
```


```{r}
bird_counts %>% 
  filter(is.na(total_hours)) %>% 
  group_by(year) %>% 
  summarise(distinct_species = n_distinct(species))
```

```{r}
bird_counts %>% 
  count(year)
```


```{r}
bird_counts %>% 
  ggplot(aes(x = year, weight = total_hours)) +
  geom_bar() +
  labs(title = "Number of hours spent watching or listening to birds",
       y = "# of hours",
       x = "year") +
  scale_y_continuous(labels = scales::comma_format())
```


```{r}
bird_counts %>% 
  filter(year >= 1950) %>% 
  group_by(year) %>% 
  summarise_at(vars(how_many_counted, total_hours), ~sum(.x, na.rm = T)) %>% 
  mutate(counted_by_hour = how_many_counted / total_hours) %>% 
  ggplot(aes(x = year, y = counted_by_hour)) +
  geom_line() +
  labs(title = "Number of birds counted per hour since 1950",
       x = "year",
       y = "# of birds")
  
```


# Latin name for species

```{r}
bird_counts %>% 
  distinct(species_latin) %>% 
  arrange(species_latin)
```

```{r}
bird_counts %>% 
  mutate(genus = str_split(species_latin, " ")) %>% 
  select(genus)

bird_counts %>% 
  mutate(genus = str_(species_latin, " ", n = 1))

bird_counts %>% 
  mutate(genus = str_remove(species_latin, " .*")) %>% 
  select(genus) %>% 
  count(genus, sort = T)
```


# Model each bird


https://r4ds.had.co.nz/many-models.html

```{r}
bird_counts_filtered <- bird_counts %>% 
  filter(year >= 1955)
```


```{r}
bird_counts_filtered %>% 
  ggplot(aes(x = year, y = how_many_counted, group = species)) +
  geom_line(alpha = 0.3) +
  scale_y_continuous(labels = scales::comma_format())
```

```{r}
bird_counts_filtered %>% 
  group_by(species) %>% 
  arrange(year) %>% 
  mutate(total_count = sum(how_many_counted),
         cum_count = cumsum(how_many_counted) / total_count) %>% 
  filter(total_count > 500) %>% 
  ggplot(aes(x = year, y = cum_count, group = species)) +
  geom_line(alpha = 0.3) +
  theme_void()
```




```{r}
wild_turkeys <- bird_counts %>% 
  filter(year >= 1955,
         species == "Wild Turkey")

wild_turkeys %>% 
  ggplot(aes(x = year, y = how_many_counted)) +
  geom_line() +
  geom_point() +
  labs(title = "Raw data") 


model_lm <- lm(how_many_counted ~ year, data = wild_turkeys)
broom::augment(model_lm) %>% 
  ggplot(aes(x = year, y = .fitted)) +
  geom_line() +
  geom_point()

```


```{r}
bird_counts_filtered %>% 
  group_by(species) %>% 
  summarise(hours = sum(total_hours))
```

# Ideas

Most popular bird
Most popular by type
Most popular species
Correlatins of species - most correlated
Gini index
Premier league rank chart

```{r}
bird_counts_filtered %>% 
  ggplot(aes(x = year, weight = total_hours)) +
  geom_bar()
```


```{r}
bird_counts_filtered %>% 
  filter(how_many_counted > 0) %>%
  group_by(year) %>% 
  summarise(n = n_distinct(species)) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_line() +
  labs(title = "The diversity of birds seen has increased",
       y = "# of distinct species",
       x = "year")
```

```{r}
bird_counts_2017 <- bird_counts_filtered %>% 
  filter(year == 2017)


  
  
```

```{r}
bird_counts_2017 
```

```{r}
bird_counts_filtered %>% 
  group_by(year) %>% 
  summarise_at(vars(total_hours, how_many_counted), ~sum(.x)) %>% 
  ggplot(aes(x = year, y = how_many_counted)) +
  geom_line()
```

```{r}
bird_counts_filtered %>% 
  group_by(year) %>% 
  summarise_at(vars(total_hours, how_many_counted), ~sum(.x)) %>%
  mutate(how_many_counted = how_many_counted / total_hours) %>% 
  ggplot(aes(x = year, y = how_many_counted)) +
  geom_line()
```

```{r}
bird_counts_2017
```

# Gini index 

Measurs the amount of inequality in a place. Is it different to the variance?

```{r}
bird_counts_filtered %>% 
  group_by(year) %>% 
  summarise(variance = sd(how_many_counted)) %>% 
  ggplot(aes(x = year, y = variance)) +
  geom_line() +
  geom_point()
```

Looks just like the number of birds counted.


```{r}

bird_species <- bird_counts_2017 %>% 
  filter(how_many_counted > 0) %>% 
  pull(species)

crossing(bird_species, bird_species) %>% 
  filter(bird_species != bird_species1) %>% 
  
  left_join(
    bird_counts_2017 %>% 
      select(bird_species = species,
             how_many_counted),
    by = "bird_species"
  ) %>% 
  
  left_join(
    bird_counts_2017 %>% 
      select(bird_species1 = species,
             how_many_counted),
    by = "bird_species1"
  ) %>% 
  
  mutate(how_many_counted.x - how)

```


https://rpubs.com/ChristoforosNikolaou/footballomics-02

If all counts were equal across all birds
```{r}

egal_count <- 
  bird_counts_2017 %>% 
  # filter(how_many_counted > 0) %>% 
  mutate(egal_count = sum(how_many_counted) / n_distinct(species)) %>% 
  pull(egal_count)

bird_counts_2017 %>% 
  # filter(how_many_counted > 0) %>% 
  mutate(egal_count = sum(how_many_counted) / n_distinct(species)) %>% 
  arrange(desc(how_many_counted)) %>% 
  top_n(30, how_many_counted) %>% 
  mutate(species = fct_reorder(species, how_many_counted)) %>% 
  ggplot(aes(x = species, y = how_many_counted)) +
  geom_point() +
  geom_hline(yintercept = egal_count) +
  coord_flip()
```


```{r}
bird_counts_2017 %>% 
  arrange(how_many_counted) %>% 
  mutate(rank_asc = row_number()) %>% 
  mutate(egal_count = sum(how_many_counted) / n_distinct(species),
         egal_count = egal_count / sum(egal_count),
         egal_count = cumsum(egal_count),
         cum_count = how_many_counted / sum(how_many_counted),
         cum_count = cumsum(cum_count)) %>% 
  select(rank_asc, ends_with("count")) %>% 
  gather(key = count_type, value = count, - rank_asc) %>% 
  ggplot(aes(x = rank_asc, y = count, group = count_type, colour = count_type)) +
  geom_line() +
  labs(title = "The most common 50 species accounted for all most all the birds seen in 2017",
       x = "rank of birds from lowest to highest number seen",
       y = " cumulative share of birds seen") +
  scale_y_continuous(labels = scales::percent_format()) +
  theme(legend.position = "none")
```


```{r}
bird_counts_2017 %>% 
  arrange(how_many_counted) %>% 
  mutate(rank_asc = row_number()) %>% 
  mutate(egal_count = sum(how_many_counted) / n_distinct(species),
         egal_count = egal_count / sum(egal_count),
         egal_count = cumsum(egal_count),
         cum_count = how_many_counted / sum(how_many_counted),
         cum_count = cumsum(cum_count)) %>% 
  select(rank_asc, ends_with("count")) %>% 
  summarise(gini = sum(egal_count - cum_count) / max(rank_asc))
```

Do the same for all the birds

```{r}
bird_counts_filtered %>% 
  group_by(species) %>% 
  summarise(how_many_counted = sum(how_many_counted)) %>% 
  arrange(how_many_counted) %>% 
  mutate(rank_asc = row_number()) %>% 
  mutate(egal_count = sum(how_many_counted) / n_distinct(species),
         egal_count = egal_count / sum(egal_count),
         egal_count = cumsum(egal_count),
         cum_count = how_many_counted / sum(how_many_counted),
         cum_count = cumsum(cum_count)) %>% 
  select(rank_asc, ends_with("count")) %>% 
  summarise(gini = sum(egal_count - cum_count) / max(rank_asc))
```


```{r}
get_gini <- function(.data) {
  
  .data %>% 
    group_by(species) %>% 
    summarise(how_many_counted = sum(how_many_counted)) %>% 
    # filter(how_many_counted > 0) %>% 
    arrange(how_many_counted) %>% 
    mutate(rank_asc = row_number()) %>% 
    mutate(egal_count = sum(how_many_counted) / n_distinct(species),
           egal_count = egal_count / sum(egal_count),
           egal_count = cumsum(egal_count),
           cum_count = how_many_counted / sum(how_many_counted),
           cum_count = cumsum(cum_count)) %>% 
    select(rank_asc, ends_with("count")) %>% 
    summarise(gini = sum(egal_count - cum_count) / max(rank_asc)) %>% 
    pull(gini)
  
}
```

```{r}
bird_counts_filtered %>% 
  group_by(year) %>% 
  nest() %>% 
  mutate(gini = map_dbl(data, ~ get_gini(.x))) %>% 
  ggplot(aes(x = year, y = gini)) +
  geom_line() +
  geom_point() +
  labs(title = "Diversity of species has increased",
       subtitle = "Gini index measures the amount of equality in number of birds seen by species.\nA value of 0 incidcates equal bird counts per species while a value of 1 indicates\nall birds seen were from a single specie",
       y = "gini index",
       x = NULL) +
  theme(
    plot.title = element_text(size = 22,
                              face = "bold")
  )
```



